---
title: "AutomaticDataExploring"
author: m00nlight <dot_wangyushi@yeah.net>
output: html_document
---

This R markdown script is use to do general data exploring. It is inspired by the script 
in a kaggle competition which can be found here [here](https://www.kaggle.com/darraghdog/springleaf-marketing-response/explore-springleaf).

This script is intended to be a general data exploring script which can be worked for any
input data.

This script depends on R packages [corrplot](https://cran.r-project.org/web/packages/corrplot/index.html), [ggplot2](http://ggplot2.org), [readr](https://cran.r-project.org/web/packages/readr/index.html), install these 
packages first to run this script.

```{r, echo = FALSE}
# load necessary library
library(corrplot)
library(ggplot2)
library(grid)
```

# Load the data you want to explore first

```{r}
data = read.csv("./input-data/boston.csv")
str(data)
```

```{r, echo = FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

# Exploring category like data

```{r, echo = FALSE}
# drop columns that is almost unique to each observation, these columns are likely 
# to be rowid etc.
delete_columns = sapply(data, function(x) length(unique(x)) >= 0.95 * nrow(data))
data = data[, !delete_columns]

for (c in colnames(data)) {
  if (length(unique(data[, c])) <= 30) {
    print(ggplot(data, aes(x = factor(1), fill = factor(data[, c]))) + geom_bar(width = 1) 
          + coord_polar(theta = "y") + ggtitle(paste("pie chart for", c)))
  }
}

```

# Exploring numeric like data

```{r, echo = FALSE}
for (c in colnames(data)) {
  tmp = unique(data[, c][!is.na(data[, c])])
  # select the column that like to be a distribution
  if(length(tmp) >= 30 && length(tmp) <= 100) {
    h = qplot(data[, c], xlab = c, ylab = "count", main = paste("histogram of", c)) +
        geom_histogram(colour="black", fill="grey") +
        theme(axis.text.x = element_text(angle = -45, hjust = 1))
    b = qplot(sample = data[, c], stat = "qq", main = paste("quantile quantile", c))
    multiplot(h, b, cols = 2)
  }
}
```

# Exploring relation between data column
```{r, echo = FALSE}
# TODO
```

# Exploring correlation of data fields

```{r, echo = FALSE}
# change data to numeric for correlation drawing.
data = data.frame(data.matrix(data))
missing_value = min(data, na.rm = TRUE)
data[is.na(data)] = missing_value - 10000
corrplot(cor(data))
```